Well-quasi-orderings on word languages

### 2025-10-16 src/appendix.tex
extra bits of information in lemma appendix

\begin{lemma}
  \label{dwclosed-wqo-complexity-width-finite:lem}
  Let $L$ be a \kl{downwards closed} language that is
  \kl{well-quasi-ordered} by the \kl{infix relation}.
  Then, the following are equivalent:
  \begin{enumerate}
    \item $L$ has bounded \kl{factor complexity},
      \label{dwclosed-wqo-complexity-width-finite:item1}
    \item $L$ has finite \kl{ordinal width}.
      \label{dwclosed-wqo-complexity-width-finite:item2}
    \item $L$ is a finite union 
      of languages of the form
      $\infset{w}$ where $w$
      is an \kl{ultimately periodic} word.
      \label{dwclosed-wqo-complexity-width-finite:item3}
  \end{enumerate}
\end{lemma}
\begin{proof}
  \cref{dwclosed-wqo-complexity-width-finite:item3} $\Rightarrow$
  \cref{dwclosed-wqo-complexity-width-finite:item1} is clear because
  \kl{ultimately periodic} words have bounded \kl{factor complexity}.

  \cref{dwclosed-wqo-complexity-width-finite:item1} $\Rightarrow$
  \cref{dwclosed-wqo-complexity-width-finite:item2} is also clear because
  unbounded \kl{factor complexity} implies the existence of arbitrarily
  large \kl{antichains}.

>>>>

\begin{lemma}
  \label{dwclosed-wqo-complexity-width-finite:lem}
  Let $L$ be a \kl{downwards closed} language that is
  \kl{well-quasi-ordered} by the \kl{infix relation}.
  Then, the following are equivalent:
  \begin{enumerate}
    \item $L$ has bounded \kl{factor complexity},
      \label{dwclosed-wqo-complexity-width-finite:item1}
    \item $L$ has finite \kl{ordinal width},
      \label{dwclosed-wqo-complexity-width-finite:item2}
    \item $L$ is a finite union of \kl{chains},
      \label{dwclosed-wqo-complexity-width-finite:item4}
    \item $L$ is a finite union 
      of languages of the form
      $\infset{w}$ where $w$
      is an ultimately periodic word.
      \label{dwclosed-wqo-complexity-width-finite:item3}
  \end{enumerate}
\end{lemma}
\begin{proof}
  First, \cref{dwclosed-wqo-complexity-width-finite:item4} $\iff$
  \cref{dwclosed-wqo-complexity-width-finite:item2} is a standard
  fact regarding \kl{ordinal width}.

  Then, \cref{dwclosed-wqo-complexity-width-finite:item3} $\Rightarrow$
  \cref{dwclosed-wqo-complexity-width-finite:item1} is clear because
  ultimately periodic words have bounded \kl{factor complexity}.

  In turn, \cref{dwclosed-wqo-complexity-width-finite:item1} $\Rightarrow$
  \cref{dwclosed-wqo-complexity-width-finite:item2} is also clear because
  unbounded \kl{factor complexity} implies the existence of arbitrarily
  large \kl{antichains}.
